---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# survParamSim

[![Travis build status](https://travis-ci.org/yoshidk6/survParamSim.svg?branch=master)](https://travis-ci.org/yoshidk6/survParamSim)

The goal of survParamSim is to perform survival simulation with parametric survival model generated from 'survreg' function in 'survival' package. 
In each simulation, coefficients are resampled from variance-covariance matrix of parameter estimates, in order to capture uncertainty in model parameters.


## Installation

You can install the package from CRAN.

``` r
install.packages("survParamSim")
```

## Example

[This GitHub pages](https://yoshidk6.github.io/survParamSim) contains function references and [vignette](https://yoshidk6.github.io/survParamSim/articles/survParamSim.html).
The example below is a sneak peek of example outputs. 

First, run survreg to fit parametric survival model:

```{r prep}
library(dplyr)
library(ggplot2)
library(survival)
library(survParamSim)

set.seed(12345)

# ref for dataset https://vincentarelbundock.github.io/Rdatasets/doc/survival/colon.html
colon2 <- 
  as_tibble(colon) %>% 
  # recurrence only and not including Lev alone arm
  filter(rx != "Lev",
         etype == 1) %>% 
  # Same definition as Lin et al, 1994
  mutate(rx = factor(rx, levels = c("Obs", "Lev+5FU")),
         depth = as.numeric(extent <= 2))
```
  
```{r fit}
fit.colon <- survreg(Surv(time, status) ~ rx + node4 + depth, 
                     data = colon2, dist = "lognormal")
```

Next, run parametric bootstrap simulation:

```{r sim}
sim <- 
  surv_param_sim(object = fit.colon, newdata = colon2, 
                 censor.dur = c(1800, 3000),
                 # Simulating only 100 times to make the example go fast
                 n.rep = 100)

sim
```

Calculate survival curves with prediction intervals:

```{r km_pi_group}
km.pi <- calc_km_pi(sim, trt = "rx", group = c("node4", "depth"))

km.pi
plot_km_pi(km.pi) +
  theme(legend.position = "bottom") +
  labs(y = "Recurrence free rate") +
  expand_limits(y = 0)
```

Calculate hazard ratios with prediction intervals:

```{r hr_pi}
hr.pi <- calc_hr_pi(sim, trt = "rx", group = c("depth"))

hr.pi
plot_hr_pi(hr.pi)
```


