---
title: "survParamSim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{survParamSim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dplyr)
library(ggplot2)
library(survival)
library(survParamSim)

set.seed(12345)
```

## Fit model with `survreg()` function

Before running a parametric survival simulation, you need to fit a model to your data using `survreg()` function of `survival` package.  
In this vignette, we will be using `colon` dataset available in `survival` package, where the treatment effect of adjuvant Levamisole+5-FU for colon cancer over placebo is evaluated.

First, we load the data and do some data wrangling.

```{r prep}
# ref for dataset https://vincentarelbundock.github.io/Rdatasets/doc/survival/colon.html
colon2 <- 
  as_tibble(colon) %>% 
  # recurrence only and not including Lev alone arm
  filter(rx != "Lev",
         etype == 1) %>% 
  # Same definition as Lin et al, 1994
  mutate(rx = factor(rx, levels = c("Obs", "Lev+5FU")),
         depth = as.numeric(extent <= 2))
```

Generating Kaplan-Meier curves for visually checking the data.  
The second plot is looking at how many censoring we have over time.  
Looks like we have a fairly uniform censoring between 1800 to 3000 days.

```{r plot_raw_data}
survfit.colon <- survfit(Surv(time, status) ~ rx, data = colon2)
survminer::ggsurvplot(survfit.colon)

survfit.colon.censor <- survfit(Surv(time, 1-status) ~ rx, data = colon2)
survminer::ggsurvplot(survfit.colon.censor)
```


Next we fit a lognormal parametric model for the data.  
Here we are using `node4` and `depth` as additional covariates in addition to treatment (`rx`).  
You can see that all of the factor has strong association with the outcome.

```{r fit}
fit.colon <- survreg(Surv(time, status) ~ rx + node4 + depth, 
                     data = colon2, dist = "lognormal")

summary(fit.colon)
```

## Perform simulation

`surv_param_sim()` is the main function of the package that takes `survreg` object as described above.  
It also require you to supply `newdata`, which is required even if it is not new - i.e. the same data was used for both `survreg()` and `surv_param_sim()`.

What it does is:
1. Re-sample all the coefficients in the parametric survival model from variance-covariance matrix for `n.rep` times.
2. Perform survival time for all subjects in `newdata` with the corresponding covariates, using one of the resampled coefficients. Also generate censoring time according to `censor.dur` (if not NULL), and replace the simulated survival time above if censoring time is earlier.
4. Repeat the steps 2. for `n.rep` times.

```{r sim}
sim <- 
  surv_param_sim(object = fit.colon, newdata = colon2, 
                 # Simulate censoring according to the plot above
                 censor.dur = c(1800, 3000),
                 # Simulate only 100 times to make the example go fast
                 n.rep = 100)
```

After the simulation is performed, you can either extract raw simulation results or further calculate Kaplan-Meier estimates or hazard ratio of treatment effect, as you see when you type `sim` in the console.

```{r simout}
sim
```

## Survival time profile

To calculate survival curves for each simulated dataset, `calc_km_pi()` can be used on the simulated object above.

```{r km_pi_calc}
km.pi <- calc_km_pi(sim, trt = "rx")
km.pi
```

Similar to the raw simulated object, you can have a few options for further processing - one of them is plotting prediction intervals with `plot_km_pi()` function.

```{r km_pi_plot}
km.pi
plot_km_pi(km.pi) +
  theme(legend.position = "bottom") +
  labs(y = "Recurrence free rate") +
  expand_limits(y = 0)
```

Or providing median survival summary table with `extract_median_surv()` function.

```{r km_pi_table}
extract_median_surv(km.pi)
```

Plot can also be made for subgroups.  
You can see that prediction interval is wide for (depth: 1 & nodes4: 1) group, mainly due to small number of subjects

```{r km_pi_group}
km.pi <- calc_km_pi(sim, trt = "rx", group = c("node4", "depth"))

plot_km_pi(km.pi) +
  theme(legend.position = "bottom") +
  labs(y = "Recurrence free rate") +
  expand_limits(y = 0)
```

## Hazard ratios (HRs)

To calculate prediction intervals of HRs, `calc_hr_pi()` can be used on the simulated object above.
Here I only generated subgroups based on "depth", since the very small N in (depth: 1 & nodes4: 1) can cause issue with calculating HRs.

```{r hr_pi}
hr.pi <- calc_hr_pi(sim, trt = "rx", group = c("depth"))

hr.pi
plot_hr_pi(hr.pi)
```


You can also extract prediction intervals and observed HR with `extract_hr_pi()` function.

```{r hr_pi_table}
extract_hr_pi(hr.pi)
```

